---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

# iSEEindex

<!-- badges: start -->
[![GitHub issues](https://img.shields.io/github/issues/iSEE/iSEEindex)](https://github.com/iSEE/iSEEindex/issues)
[![GitHub pulls](https://img.shields.io/github/issues-pr/iSEE/iSEEindex)](https://github.com/iSEE/iSEEindex/pulls)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check-bioc](https://github.com/iSEE/iSEEindex/workflows/R-CMD-check-bioc/badge.svg)](https://github.com/iSEE/iSEEindex/actions)
[![Codecov test coverage](https://codecov.io/gh/iSEE/iSEEindex/branch/main/graph/badge.svg)](https://app.codecov.io/gh/iSEE/iSEEindex?branch=main)
<!-- badges: end -->

The goal of `iSEEindex` is to provide an interface to any collection of data sets within an iSEE web-application.

The main functionality of this package is to define a custom landing page allowing app maintainers to list a custom collection of data sets that users can selected from and directly load objects into an iSEE web-application.

## Installation instructions

Get the latest stable `R` release from [CRAN](http://cran.r-project.org/). Then install `iSEEindex` from [Bioconductor](http://bioconductor.org/) using the following code:

```{r 'install', eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}

BiocManager::install("iSEEindex")
```

And the development version from [GitHub](https://github.com/iSEE/iSEEindex) with:

```{r 'install_dev', eval = FALSE}
BiocManager::install("iSEE/iSEEindex")
```

## Example

This is a basic example which shows you how to launch an application that lists publicly available data sets hosted on [zenodo.org][zenodo-repository]:

```{r example, eval = requireNamespace('iSEEindex'), message=FALSE, warning=FALSE}
library("iSEEindex")
library("BiocFileCache")

bfc <- BiocFileCache()

dataset_fun <- function() {
    out <- read.csv(system.file(package = "iSEEindex", "datasets.csv"))
}

app <- iSEEindex(bfc, dataset_fun)

if (interactive()) {
  shiny::runApp(app, port = 1234)
}
```

## Citation

Below is the citation output from using `citation('iSEEindex')` in R. Please
run this yourself to check for any updates on how to cite __iSEEindex__.

```{r 'citation', eval = requireNamespace('iSEEindex')}
print(citation('iSEEindex'), bibtex = TRUE)
```

Please note that the `iSEEindex` was only made possible thanks to many other R and bioinformatics software authors, which are cited either in the vignettes and/or the paper(s) describing this package.

## Code of Conduct

Please note that the `iSEEindex` project is released with a [Contributor Code of Conduct](http://bioconductor.org/about/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

## Development tools

* Continuous code testing is possible thanks to [GitHub actions](https://www.tidyverse.org/blog/2020/04/usethis-1-6-0/)  through `r BiocStyle::CRANpkg('usethis')`, `r BiocStyle::CRANpkg('remotes')`, and `r BiocStyle::CRANpkg('rcmdcheck')` customized to use [Bioconductor's docker containers](https://www.bioconductor.org/help/docker/) and `r BiocStyle::Biocpkg('BiocCheck')`.
* Code coverage assessment is possible thanks to [codecov](https://codecov.io/gh) and `r BiocStyle::CRANpkg('covr')`.
* The [documentation website](http://isee.github.io/iSEEindex) is automatically updated thanks to `r BiocStyle::CRANpkg('pkgdown')`.
* The code is styled automatically thanks to `r BiocStyle::CRANpkg('styler')`.
* The documentation is formatted thanks to `r BiocStyle::CRANpkg('devtools')` and `r BiocStyle::CRANpkg('roxygen2')`.

For more details, check the `dev` directory.

This package was developed using `r BiocStyle::Biocpkg('biocthis')`.

## Code of Conduct

Please note that the iSEEindex project is released with a [Contributor Code of Conduct](http://bioconductor.org/about/code-of-conduct/). By contributing to this project, you agree to abide by its terms.

<!-- Links -->

[zenodo-repository]: https://zenodo.org/record/7304331
